---
title: "NSRL19A Females Weights"
author: "Frederico Kiffer, PhD"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

Import Libraries
```{r}
library(tidyverse)
library(colorspace)
library(ggpubr)
library(rstatix)
```

# View and Format Data

Load and View Data
```{r}
FemalesWeights <- read_csv('https://raw.githubusercontent.com/FKiffer/NSRL19AFemales/main/Female%20Weights.csv')
FemalesWeights
```

Define Factor Variables
```{r}
FemalesWeights <- FemalesWeights %>%
  convert_as_factor(Radiation, Diet, Mouse)
```

Drop Final Weights Collected for Normal Distribution
```{r}
FemalesWeightsClean <- FemalesWeights %>%
  filter(Date < "2020-04-24")
```

Mean and SEM by Day
```{r}
FemalesWeightsClean %>%
  group_by(Radiation, Diet, Date) %>%
  get_summary_stats(Weight, type = "mean_se")
```

View Raw Data
```{r}
ggplot(FemalesWeightsClean, aes(x = Date, y = Weight,
  shape = Diet, color = Radiation, fill = Diet, 
  group = interaction(Radiation, Diet))) +
  geom_point() +
  scale_color_manual(values = c("#db8270", "#3e8dcc")) +
  ggtitle("19A Females Weights") + 
  ylab("Weight (g)") + 
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1), 
      panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
      plot.title = element_text(hjust = 0.5, face="bold", size = 18), 
      panel.background = element_blank(), 
      axis.line = element_line(colour = "black", size = 1), 
      axis.text.y=element_text(colour = "black", family="Arial", size=12,
                             face="bold"), 
      axis.title.x=element_text(face = "bold", size=14),
      axis.title.y=element_text(face = "bold", size=14),
      legend.title=element_text(face="bold"),
      legend.text=element_text(face="bold"))
```

# Test for ANOVA Assumptions

Identify Outliers
```{r}
FemalesWeightsCleanOutliers <- FemalesWeightsClean %>%
  group_by(Radiation, Diet, Date) %>%
  identify_outliers(Weight)
FemalesWeightsCleanOutliers
```

Test for Normality of Distributions at Each Date
```{r}
FemalesWeightsClean %>%
  group_by(Radiation, Diet, Date) %>%
  shapiro_test(Weight)
```

Visualize Distributions by QQPlots by Increasing Dates
```{r fig.align="center", fig.width = 8}
FemalesClean1 <- FemalesWeightsClean %>% filter(Date <= '2019-07-31')
FemalesClean2 <- FemalesWeightsClean %>% filter(Date > '2019-07-31' & 
                                                  Date <= '2019-10-30')
FemalesClean3 <- FemalesWeightsClean %>% filter(Date > '2019-10-30' &
                                                  Date <= '2020-01-08')
FemalesClean4 <- FemalesWeightsClean %>% filter(Date > '2020-01-08')

ggqqplot(FemalesClean1, "Weight", ggtheme = theme_bw()) +
  facet_grid(Radiation + Diet ~ Date,
             labeller = "label_both")
ggqqplot(FemalesClean2, "Weight", ggtheme = theme_bw()) +
  facet_grid(Radiation + Diet ~ Date,
             labeller = "label_both")
ggqqplot(FemalesClean3, "Weight", ggtheme = theme_bw()) +
  facet_grid(Radiation + Diet ~ Date,
             labeller = "label_both")
ggqqplot(FemalesClean4, "Weight", ggtheme = theme_bw()) +
  facet_grid(Radiation + Diet ~ Date,
             labeller = "label_both")
```

Test for Homogeneity of Variance
```{r}
HomVar <- FemalesWeightsClean %>%
  group_by(Date) 

HomVarRad <- HomVar %>%
  levene_test(Weight ~ Radiation)

HomVarDiet <- HomVar %>%
  levene_test(Weight ~ Diet)

HomVarRad
HomVarDiet
```
# Repeated Measures 3-Way ANOVA

Repeated Measures ANOVA with Greenhouse-Geisser Corrections for Spherecity
```{r}
FemalesANOVA <- anova_test(
  data = FemalesWeightsClean, Weight ~ Radiation * Diet * Date, wid = Mouse,
  within = Date
)
get_anova_table(FemalesANOVA)
```
Because our data contains missing samples due to animal deaths, a mixed model is preferred:

Repeated Measures 3-Way Mixed Model ANOVA
```{r}
FemalesMMANOVA <- anova_test(
  data = FemalesWeightsClean, dv = Weight, between = c(Radiation, Diet), 
  within = Date, wid = Mouse
)
get_anova_table(FemalesMMANOVA)
```

# Multiple Comparisons

Here are several approaches to multiple comparisons that use different pooling Standard Deviations (SD), depending on which main variables are relevant for comparisons

Multiple Comparisons with Bonferroni Corrections by Diet (pools all Diet SDs)
```{r}
PairwiseDiet <- FemalesWeightsClean %>%
  group_by(Date, Radiation) %>%
  pairwise_t_test(
    Weight ~ Diet, paired = FALSE,
    p.adjust.method = "bonferroni"
  )
PairwiseDiet
```

Multiple Comparisons with Bonferroni Corrections by Radiation (pools all Radiation SDs)
```{r}
PairwiseRad <- FemalesWeightsClean %>%
  group_by(Date, Diet) %>%
  pairwise_t_test(
    Weight ~ Radiation, paired = FALSE,
    p.adjust.method = "bonferroni"
  )
PairwiseRad
```

Format Data for Multi-Variable Multiple Comparisons
```{r}
FemalesWeightsClean$Group <- paste(FemalesWeightsClean$Diet, 
                                   FemalesWeightsClean$Radiation)
head(FemalesWeightsClean)
```

Multiple Comparisons with Bonferroni Corrections Across Diet and Radiation (pools all SDs)
```{r}
PairwiseAll <- FemalesWeightsClean %>%
  group_by(Date) %>%
  pairwise_t_test(
    Weight ~ Group, paired = FALSE,
    p.adjust.method = "bonferroni"
  )
PairwiseAll
```

Multiple Comparisons with Bonferroni Corrections Across Diet and Radiation (Does not pool SDs)
```{r}
PairwiseAll <- FemalesWeightsClean %>%
  group_by(Date) %>%
  t_test(
    Weight ~ Group, paired = FALSE,
    p.adjust.method = "bonferroni"
  )
PairwiseAll
```

# Effect Sizes

Main Effect Size (Eta Squared)
```{r}
FemaleWeightsES <- FemalesWeightsClean %>%
  group_by(Date) %>%
  convert_as_factor(Date)
eta_squared(aov(Weight ~ Radiation * Diet, data = FemalesWeightsClean))
```

Multiple Comparisons Effect Size (Cohen's d)
```{r}
cohens_d(Weight ~ Group, data = FemaleWeightsES)
```

# Figure

Smoothened Line Graph (Mean +/- SEM)
```{r}
FemalesWeightsGraph <- ggplot(FemalesWeightsClean, aes(x=Date, 
    y=Weight, group = interaction(Radiation, Diet), 
    fill=Diet, color=Radiation)) +
    geom_point(size = 1, alpha = 0) +
    geom_smooth(span = .4, aes(color=Radiation, fill=Diet)) +
   scale_fill_manual(values = c( "#CACED4", "#FBF293"))+
    scale_color_manual(values = c("#db8270", "#3e8dcc"))+
  ggtitle("Females Weights") + 
  ylab("Weight (g)") + 
  xlab("Date") +
  scale_y_continuous(expand = c(0,0)) +
  #scale_x_continuous(expand = expansion(mult = c(0,0))) +
  #makes x axis start at 1 if scaled continuously
  coord_cartesian(expand = FALSE, ylim = c(20,40)) + 
  #does not remove points outside of y-limit, merges y-axis with line graph
  scale_x_date(date_labels = "%b %d", date_breaks = "1 month") +
  #scales x axis by date, every month
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1),
      plot.title = element_text(hjust = 0.5, face="bold", size = 18), 
      panel.background = element_blank(), 
      axis.line = element_line(colour = "black", size = 1), 
      axis.text=element_text(colour = "black", family="Arial", size=12,
                             face="bold"), 
      axis.title.x=element_text(face = "bold", size=14),
      axis.title.y=element_text(face = "bold", size=14),
      legend.title=element_text(face="bold"),
      legend.text=element_text(face="bold")) 
print(FemalesWeightsGraph)
```

To Save as SVG:
```{r}
#library(svglite)
#ggsave("FemalesWeightsGraph.svg", width = 8, height = 4)
```