---
title: "NSRL 19A Females Survival"
author: "Frederico Kiffer, Ph.D."
output:
  html_document:
    df_print: paged
  pdf_document: default
---

Here we conduct survival analyses of female mice following a 75cGy whole-body exposure to NASA's ground-based 33-beam Galactic Cosmic Ray analog or sham-irradiation during the 19A campaign. Mice also received the candidate countermeasure CDDO-EA or just a vehicle formulation.

Load Libraries
```{r}
library(colorspace)
library(survival)
library(survminer)
library(ggpubr)
```

Read CSV
```{r}
FemalesSurvival <- read.csv("https://raw.githubusercontent.com/FKiffer/NSRL19AFemales/main/FemalesSurvival.csv")
FemalesSurvival
```

Convert Variables as Factors
```{r}
cols <- c("Vehicle", "CDDO", "Sham", "GCR")
FemalesSurvival[cols] <- lapply(FemalesSurvival[cols], factor)
head(FemalesSurvival)
```


Designate Variables
```{r}
km.females_survival <- survfit(Surv(
  time = FemalesSurvival$Weeks,
  event = FemalesSurvival$Death) ~ Vehicle+CDDO+Sham+GCR,
  data = FemalesSurvival, type = "kaplan-meier"
  )
summary(km.females_survival)
```

Log-Rank Test
```{r}
survdiff(Surv(
  time = FemalesSurvival$Weeks,
  event = FemalesSurvival$Death) ~ Vehicle+CDDO+Sham+GCR,
  data = FemalesSurvival)
```

Graph with 95% Confidence Interval
```{r}
FemalesSurvivalGraph <- ggsurvplot(km.females_survival, conf.int = TRUE, 
    pval = FALSE, 
    palette = c("#ef5e4f", "#4F90EF", "#ef5e4f", "#4F90EF"),
    risk.table = FALSE,
    legend.labs = c("CDDO/GCR", "CDDO/Sham", "Veh/GCR", "Veh/Sham"), 
    legend.title = "Diet/Radiation", legend = c("right"), break.time.by = 5, 
    axes.offset = FALSE, ylim = c(0, 1), linetype=c("strata"),
    xlim = c(0, 60), ggtheme = theme_classic() + #ad breaks
    theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 18), 
          axis.line = element_line(colour = "black", size = 1),
          axis.text=element_text(colour = "black", family = "Arial", 
            size = 12, face="bold"), 
          legend.title = element_text(face = "bold", size = 10),
          legend.text = element_text(face = "bold", size = 8)),
    font.x = c(14, "bold"), font.y = c(14, "bold"),
    font.legend.title = c(10, "bold")) +
    ggtitle("Mouse Attrition") +
    ylab("Mouse Survival Probability") +
    xlab("Weeks From Irradiation") 
FemalesSurvivalGraph
```

Save as SVG (optional)
```{r}
#library(svglite)
#ggsave("19AFemalesSurvival.svg", width = 8, height = 4)
```

Cox Proportional Hazards
```{r}
CoxPH <- coxph( Surv(
  time = FemalesSurvival$Weeks,
  event = FemalesSurvival$Death) ~ Vehicle+CDDO+Sham+GCR, #(No animals receiving veh died)
  data = FemalesSurvival)
CoxPH
summary(CoxPH)
```

Cox Proportionate Hazards does not work when there are groups without deaths, and cannot appropriately counterbalance for groups that do have deaths in a 2x2 experimental design. 

